Moment Distances from Robust Subspace for Botnet Detection

### 2019-05-15 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
working on results

Mahalanobis Distance (MD) is a generalized distance which is useful for determining the similarity between an unknown sample and a collection of known samples, by considering the correlations between the variables and their mean values. MD has been used for distance based anomaly detection with robust estimates in many areas, assuming that the Mahalanobis Distance between robust estimates and new observations can reveal anomalies.

@@

In this paper, scalars are denoted by italic letters (\emph{a, b, A, B, $α$, $β$}), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}. The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively. The Frobenius norm is denoted as $\left\| \mathord{\cdot} \right\|_F$, while $\left\| \mathord{\cdot} \right\|_*$ denotes the nuclear norm of a matrix and $\left\| \mathord{\cdot} \right\|_1$ means the sum of the absolute values of matrix entries.

@@

We refer to Garcia \cite{garcia2014identifying} and Garcia \emph{et al.} \cite{garcia2014empirical} for a detailed description of the performed attacks and C\&C flows, also including the topology of the adopted testbed, rules for classifing normal flows, and an analysis of behaviors or patterns of the malware's traffic.

@@

Related works proposed some features for improving network attack detection in various scenarios and approaches\cite{lakhina2005mining,gu2008botminer,callegari2011novel,chandrashekar2014survey,garcia2014empirical,acarali2016survey}. Inpired by these proposals and based on previous experience \cite{vieira2017model, galibus2017offline} and cross validation, we propose the following features for improving the anomaly detection by means of a method based on higher moments distances from robust subspace.

@@

The proposed features are counting and average calculations, therefore it is necessary to adopt time windows to aggragate flows and calculate the features. We evaluate windows of 0.15 and 0.25 seconds, considering account that lower windows reduce the meantime between the data collection and results of the botnet detection approach, as well as permit to have more details to exactly identify malicious flows and attackers.

Figure \ref{fig:fig02} shows the distribution of average flow duration, for scenario 12, aggregated by windows of 0.15 seconds. In comparison to Figure \ref{fig:fig01}, that shows the distribution of flow duration for the same scenario, we can see that the feature based on aggration and on average flow duration is more discriminative and shows a better separation between normal and anomalous classes. Therefore, such data can improve the anomaly detection by learning algorithms.

@@

The resulting data set is modeled into the matrix $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ with $n$ rows as observations and $p$ columns of features, and is divided into scenarios, where each $i$-th scenario, presented by ID column in Table \ref{tab:tab01}, are represented as $\boldsymbol{X}_i$ and can be denoted as $\boldsymbol{X}_{18-2}$, for example. The normal traffic, required for trainning the semi-supervised approach, is denoted as $\boldsymbol{X}_i^r$

@@

This section describes the proposed approach for botnet detection by means of a disntance analysis between moments computed from a robust subspace and new observations of network traffic.

Robust subspace learning can be defined as the decomposition of a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise. We propose to learn a robust subspace $\textbf{L}$ from the normal traffic $\textbf{X}_i^r$ for computing robust moments, i.e. the mean $\boldsymbol{\bar{x}}$, skewness $\boldsymbol{\bar{s}}$ and kurtosis $\boldsymbol{\bar{k}}$, in order to evaluate the distance $\boldsymbol{d}$ of new observations $\boldsymbol{X}^t$ of these moments and classify as anomalous the observations with the largest distances.

Robust PCA is a well known method to recover a low-rank matrix $\textbf{L}$ and sparce matrix $\textbf{S}$ from corrupted measurements modeled as $\textbf{X} = \textbf{L} + \textbf{S}$. This decomposition in low-rank and sparse matrices can be achieved by techniques such as Principal Component Pursuit method (PCP), and by optimization methods, such as the Augmented Lagrange Multiplier Method (ALM), Alternating Direction Method (ADM), Fast Alternating Minimization (FAM) or Iteratively Reweighted Least Squares (IRLS) \cite{candes2011robust,vaswani2018robust,lerman2018overview}. We adotp RPCA with ALM, which can be formulated as

\begin{equation}\label{eq:eq01}
	\min_{\boldsymbol{L,S}}\left\|\boldsymbol{L}\right\|_* + \lambda\left\|\boldsymbol{S}\right\|_1 + \langle \boldsymbol{Y, X - L - S}  \rangle + \frac{\mu}{2}\left\|\boldsymbol{X - L - S}\right\|_F^2,
\end{equation}
where $\boldsymbol{Y}$ are the Lagrange multipliers and have the constraing $\boldsymbol{C = D - A - E}$...

Initially, the data $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ is split into normal data $\textbf{X}^r$ for rosbust subspace learning, and into $\textbf{X}^t$ for testing the anomaly detection on contaminated data, that can also be viewed as new observations of a production environment.

@@

\begin{equation}\label{eq:eq02}
	\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} \boldsymbol{x}_i, 
\end{equation}

@@

\begin{equation}\label{eq:eq03}
	\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime,
\end{equation}

The general formula of the $u$-th moment can be expressed as
\begin{equation}\label{eq:eq02}
	m_u = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}( x_i - \bar{x})^u.
\end{equation}

Therefore, the skewness is calculated by
\begin{equation}\label{eq:eq03}
	\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}},
\end{equation}

and the kurtosis as
\begin{equation}\label{eq:eq04}
	\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2} .
\end{equation}

We propose to compute the mean $\boldsymbol{\bar{x}}$, the skewness $\boldsymbol{\bar{s}}$, the kurtosis $\boldsymbol{\bar{k}}$ and the covariance matrix $\boldsymbol{\bar{\Sigma}}$ from $\boldsymbol{L}^r$, after to minimize the Equation \ref{eq:eq01} for $\textbf{X}^r$. We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies betweem new observations and the moments calculated from a robust subspace computed by RPCA.

The classical Mahalanobis Distance is defined as		
	\begin{equation}\label{eq:eq05}
		\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime},
	\end{equation}
where $\boldsymbol{x}$ is a vector of a new observation, $\bar{\boldsymbol{x}}$ is the mean vector of known observations, also referred as location, and $\boldsymbol{\bar{\Sigma}}$ is the covariance matrix of known observations, also referred as scatter. The classifical MD is usually relies on robust mean and robust covariace matrix for outlier detection, which are commonly computed by Fast-MCD \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}. Here we propose to calculate the MD based on mean and covariance matrix computed from a robust subspace learned by RPCA. We also propose to extend the Equation \ref{eq:eq05} to implement a Skewness-based Mahalanobis Distance, as follows:

@@

The distances shall be computed and evaluated separatelly for botnet detection, in order to verify if the scenarios have different patterns that can be explored differently by distances of each moment individually.

The Algorithmm \ref{alg:alg01} presents a overview the experiment
\begin{algorithm}
	\label{alg:alg01}
	\scriptsize
	\SetAlgoLined
	\KwResult{$\boldsymbol{\bar{t}}$}
	Given the data set $\boldsymbol{X}$ split into 13 scenarios $\boldsymbol{X}_i$, and susequently split into normal observations $\boldsymbol{X}_i^n$ and contaminated observations $\boldsymbol{X}_i^t$\;
	\For{$eacch scenario$}{
		\If{$\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ exist}{
			Put $\boldsymbol{\bar{x}}_1 = \boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_1 = \boldsymbol{\hat{S}}_2$\;
		}
		Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,..., n$\;
		Sort $\boldsymbol{d}_1( i )$, which yields an index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
		Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), ..., \pi(h)\}$\;
		Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
		Compute $det(\boldsymbol{\hat{S}}_2)$ and $det(\boldsymbol{\hat{S}}_1)$\;			
	}
	Compute $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$ for last $\boldsymbol{h}_2$\;
	\caption{C-Step with higher moments}
\end{algorithm}

The expected result $\boldsymbol{\bar{t}}$ refers to the predicted classification of the new observations as normal or anomalous.

Only for evaluation pourpose, we use part of the contaminated data, denoted as $\textbf{X}^v$, for cross-validation, in order to evaluate the contamination rate $c$ of the target data. The contamination rate parameter is widely adopted for well stablished outlier detection algorithms \cite{zhao2019pyod}, and refers to the percentage of observations that are classified as anomalous, which can be well known for some areas, or can be computed by cross-validation or can be assumed according to previous observations.

Therefore we have $\textbf{L}$...

Fast-MCD is commonly used as \textbf{supervised} or \textbf{unsupervised} algorithm for anomaly detection.

\textbf{Supervised} approach:

Compute Fast-MCD for training data with known normal and anomalous data.
Select the best threshold $t$ for detecting known anomalies according to largest distances $\boldsymbol{d}(i)$.
Save the best $t, \boldsymbol{\bar{x}}$ and $\boldsymbol{\hat{S}}$.
Compute $d(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\hat{S}})$ for new observations.
Classify observations with $\boldsymbol{d}(i)$ higher than $t$ as anomalous.

\textbf{Unsupervised} approach:

Given a known contamination $c$, which is the percentage of anomalies in a data set.
Compute Fast-MCD and $d(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\hat{S}})$ for new observations.
Classify the $c$ observations with largest distances as anomalous.

Finally, We propose the following \textbf{Semi Supervised} approach for Fast-MCD based anomaly detection:

Compute Fast-MCD for training data containing only known normal observations.
Save the fitted $\boldsymbol{\bar{x}}_{normal}$ and $\boldsymbol{\hat{S}}_{normal}$.
Carry out cross validation on known contaminated data in order to identify the contamination $c$ that best classify anomalies and normal data.
Compute $d(\boldsymbol{x},\bar{\boldsymbol{x}}_{normal}, \boldsymbol{\hat{S}}_{normal})$ for new observations.
The $c$ observations with largest $\boldsymbol{d}(i)$ are classified as anomalous.

@@

Experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statistical approaches, which are K-means and Gaussian Mixture Model (GMM) \cite{gaddam2007kmeans,moustafa2019holistic}, respectively. We also evaluate the results of ROBPCA \cite{hubert2005robpca}, which is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness. In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of CTU-13 individually, in a semi-supervised approach that does not rely on training data with anomalies.

This section presents the performed experiments and the acquired results. First, in Section \ref{sec:AnalyzedScenario}, the experimental scenario adopted in the experiments is summarized. Then, Section \ref{sec:largesteigenvaluesanalysis} shows the results of the largest eigenvalue analysis by time frames for the experimental scenario. In Section \ref{sec:MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection in the simulated data set. Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, Section \ref{sec:EigenSimilarityAnalysis} shows the results of similarity analysis for detailed flood and port scan identification for the experimental scenario. Section \ref{sec:DarpaEvaluation} presents the results of the largest eigenvalue analysis, model order selection and the eigenvalue analysis for flood and probe attack detection in the DARPA 1998 data set.

@@

Given a contaminated data set, split it into normal (60\%), cross validation (20\% normal + 50\% anomalous) and test (20\% normal + 50\% anomalous).

@@

Clustering methods for anomaly detection rely on three key assumptions: legitimate data instances often fall into a cluster whereas attacks do not, legitimate data instances are usually located near the closest cluster centroid while anomaly ones are often far away from it, legitimate data instances fall into vast and dense clusters and anomalies into small or spare ones \cite{ahmed2016survey, moustafa2019holistic}. We rely on this for our implementation

GMM is a statistical-based and parametric approach algorithm that estimates the distribution of the normal class from a training set and is typically based on a set of kernels \cite{moustafa2019holistic}.

The Mahalanobis Distance is a measure of the distance between a vector $\boldsymbol{x}$ and a distribution $\boldsymbol{X}$, introduced by P. C. Mahalanobis in 1936 \cite{mahalanobis1936md}. It is a multi-dimensional generalization for measuring how many standard deviations away $\boldsymbol{x}$ is from the mean $\boldsymbol{\bar{x}}$ and covariance $\boldsymbol{\hat{S}}$ of $\boldsymbol{X}$.

@@

Gargia et al. \cite{garcia2014empirical} propose a division of CTU-13's scenarios into train and test, for evaluation of network attack detection algorithms. 
However, we decide to evaluate our proposed approach for each scenario individually, considering that in the division proposed by Gargia et al. some bots are present only for training and not for testing. 
Although our proposal does not rely on anomalous data for training, we desire to evaluate our proposal to detect all types of bots captured by CTU-13 data set.

Why GMM?

As clustering-based method, we selected K-means, which is the clustering algorithm most adopted for general purposes and for network anomaly detection \cite{gaddam2007kmeans,moustafa2019holistic}.
It is observed that network traffic do not belong to a Gaussian distribution \cite{benson2010network,moustafa2019holistic}, algorithms based on non-Gaussian distributions, such as GMM, has been adopted for network anomaly detection \cite{moustafa2019holistic}.

@@

This paper evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work \cite{tenorio2013greatest} and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection. Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.

@@

>>>>

Mahalanobis Distance (MD) is a generalized distance which is useful for determining the similarity between an unknown sample and a collection of known samples, by considering the correlations between the variables and their mean values. The MD is a measure of the distance between a vector $\boldsymbol{x}$ and a distribution $\boldsymbol{X}$, introduced by P. C. Mahalanobis in 1936 \cite{mahalanobis1936md}. It is a multi-dimensional generalization for measuring how many standard deviations away $\boldsymbol{x}$ is from the mean $\boldsymbol{\bar{x}}$ and covariance $\boldsymbol{\bar{\Sigma}}$ of $\boldsymbol{X}$. MD has been used for distance based anomaly detection with robust estimates in many areas, assuming that the Mahalanobis Distance between robust estimates and new observations can reveal anomalies.

@@

In this paper, scalars are denoted by italic letters (\emph{a, b, A, B, $α$, $β$}), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}. The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively. The Frobenius norm is denoted as $\left\| \mathord{\cdot} \right\|_F$, while $\left\| \mathord{\cdot} \right\|_*$ denotes the nuclear norm of a matrix and $\left\| \mathord{\cdot} \right\|_1$ means the sum of the absolute values of matrix entries. We define the operator $[ \mathord{\cdot} ]^n$, which denotes the $n$ largest values of a vector.

@@

We refer to Garcia \cite{garcia2014identifying} and Garcia \emph{et al.} \cite{garcia2014empirical} for a detailed description of the performed attacks and C\&C flows, also including the topology of the adopted testbed, rules for classifying normal flows, and an analysis of behaviors or patterns of the malware's traffic.

@@

Related works proposed some features for improving network attack detection in various scenarios and approaches\cite{lakhina2005mining,gu2008botminer,callegari2011novel,chandrashekar2014survey,garcia2014empirical,acarali2016survey}. Inspired by these proposals and based on previous experience \cite{vieira2017model, galibus2017offline} and cross validation, we propose the following features for improving the anomaly detection by means of a method based on higher moments distances from robust subspace.

@@

The proposed features are counting and average calculations, therefore it is necessary to adopt time windows to aggregate flows and calculate the features. We evaluate windows of 0.15 and 0.25 seconds, considering account that lower windows reduce the meantime between the data collection and results of the botnet detection approach, as well as permit to have more details to exactly identify malicious flows and attackers.

Figure \ref{fig:fig02} shows the distribution of average flow duration, for scenario 12, aggregated by windows of 0.15 seconds. In comparison to Figure \ref{fig:fig01}, that shows the distribution of flow duration for the same scenario, we can see that the feature based on aggregation and on average flow duration is more discriminative and shows a better separation between normal and anomalous classes. Therefore, such data can improve the anomaly detection by learning algorithms.

@@

The resulting data set is modeled into the matrix $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ with $n$ rows as observations and $p$ columns of features, and is divided into scenarios, where each $i$-th scenario, presented by ID column in Table \ref{tab:tab01}, are represented as $\boldsymbol{X}_i$ and can be denoted as $\boldsymbol{X}_{18-2}$, for example. The normal traffic, required for training the semi-supervised approach, is denoted as $\boldsymbol{X}_i^r$

@@

This section describes the proposed approach for botnet detection by means of a distance analysis between moments computed from a robust subspace and new observations of network traffic.

Robust subspace learning can be defined as the decomposition of a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise. We propose to learn a robust subspace $\textbf{L}$ from the normal traffic $\textbf{X}_i^r$ for computing $\textbf{L}_i^r$, $\textbf{S}_i^r$ and the robust moments, i.e. the mean $\boldsymbol{\bar{x}}$, skewness $\boldsymbol{\bar{s}}$ and kurtosis $\boldsymbol{\bar{k}}$, in order to evaluate the distance $\boldsymbol{d}$ of new observations $\boldsymbol{X}^t$ to these moments and classify as anomalous the observations with the largest distances.

Robust PCA is a well known method to recover a low-rank matrix $\textbf{L}$ and sparce matrix $\textbf{S}$ from corrupted measurements modeled as $\textbf{X} = \textbf{L} + \textbf{S}$. This decomposition in low-rank and sparse matrices can be achieved by techniques such as Principal Component Pursuit method (PCP), and by optimization methods, such as the Augmented Lagrange Multiplier Method (ALM), Alternating Direction Method (ADM), Fast Alternating Minimization (FAM) or Iteratively Reweighted Least Squares (IRLS) \cite{candes2011robust,vaswani2018robust,lerman2018overview}. We adopt RPCA with ALM, which can be formulated as

\begin{equation}\label{eq:eq01}
	\min_{\boldsymbol{L,S}}\left\|\boldsymbol{L}\right\|_* + \lambda\left\|\boldsymbol{S}\right\|_1 + \langle \boldsymbol{Y, X - L - S}  \rangle + \frac{\mu}{2}\left\|\boldsymbol{X - L - S}\right\|_F^2,
\end{equation}
where $\boldsymbol{Y}$ are the Lagrange multipliers and have the constraint $\boldsymbol{C = D - A - E}$...

@@

\begin{equation}\label{eq:eq02}
	\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} \boldsymbol{x}_i, 
\end{equation}

@@

\begin{equation}\label{eq:eq03}
	\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime,
\end{equation}

The general formula of the $u$-th moment can be expressed as
\begin{equation}\label{eq:eq04}
	m_u = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}( x_i - \bar{x})^u.
\end{equation}

Therefore, the skewness is calculated by
\begin{equation}\label{eq:eq05}
	\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}},
\end{equation}

and the kurtosis as
\begin{equation}\label{eq:eq06}
	\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}.
\end{equation}

We propose to compute the mean $\boldsymbol{\bar{x}}$, the skewness $\boldsymbol{\bar{s}}$, the kurtosis $\boldsymbol{\bar{k}}$ and the covariance matrix $\boldsymbol{\bar{\Sigma}}$ from the robust subspace $\boldsymbol{L}^r$, after to minimize the Equation \ref{eq:eq01} for $\textbf{X}^r$. We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies between new observations and the moments calculated from a robust subspace computed by RPCA.

The classical Mahalanobis Distance is defined as		
\begin{equation}\label{eq:eq07}
	\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime},
\end{equation}
where $\boldsymbol{x}$ is a vector of a new observation, $\bar{\boldsymbol{x}}$ is the mean vector of known observations, also referred as location, and $\boldsymbol{\bar{\Sigma}}$ is the covariance matrix of known observations, also referred as scatter. The classical MD usually relies on robust mean and robust covariance matrix for outlier detection, which are commonly computed by Fast-MCD \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}. Here we propose to calculate the MD based on mean and covariance matrix computed from a robust subspace learned by RPCA. We also propose to extend the Equation \ref{eq:eq05} to implement a Skewness-based Mahalanobis Distance, as follows:

@@

The distances $\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}})$, $\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}})$ and $\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}})$ shall be computed and evaluated separately for botnet detection. Therefore, the robust subspace learning from normal data combined to MD can be called md-RPCA when using mean-based MD, sd-RPCA when using skewness-based MD, and kd-RPCA when using kurtosis-based MD.

The contamination rate parameter is widely adopted for well established outlier detection algorithms \cite{zhao2019pyod}, and refers to the percentage of observations that are classified as anomalous. The contamination rate can be well known for some areas, or can be computed by cross-validation or can be assumed according to previous observations. In our proposal, the contamination define the rate of highest distances $\boldsymbol{d}$ that shall be classified as anomalous. Therefore, the resulting $\boldsymbol{\bar{t}}$ refers to observations classified as anomalous according to $c$.

The Algorithm \ref{alg:alg01} presents how the proposed Moment Distances from Robust Subspace for Botnet Detection can be algorithmically described.
\begin{algorithm}
	\label{alg:alg01}
	\SetAlgoLined
	\KwResult{$\boldsymbol{\bar{t}}$}
	Given $\boldsymbol{X}$ split into $\boldsymbol{X}^n$ and $\boldsymbol{X}_i^t$\;
	$\min_{\boldsymbol{L}^n, \boldsymbol{S}^n}\left\|\boldsymbol{L}^n\right\|_* + \lambda\left\|\boldsymbol{S}^n\right\|_1 + \langle \boldsymbol{Y}^n, \boldsymbol{X}^n - \boldsymbol{L}^n - \boldsymbol{S}^n  \rangle + \frac{\mu}{2}\left\|\boldsymbol{X}^n - \boldsymbol{L}^n - \boldsymbol{S}^n\right\|_F^2$\;
	$\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} \boldsymbol{x}_i$, for $\boldsymbol{L}^n$\;
	$\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime$, for $\boldsymbol{L}^n$\;
	$\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}}$, for $\boldsymbol{L}^n$\;
	$\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}$, for $\boldsymbol{L}^n$\;
	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime}$\;
	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{s}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{s}})^\prime}$\;
	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{k}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{k}})^\prime}$\;
	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}})]^c$\;
	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}})]^c$\;
	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}})]^c$\;
	\caption{Moment Distances from Robust Subspace}
\end{algorithm}

The steps between 1 and 6 of the Algorithm \ref{alg:alg01} are the training from normal data $\boldsymbol{X}^n$, while the steps between 7 and 12 aim the anomaly detection on new contamianted observations. Note that the steps 7 and 10 refers to the md-RPCA, while the steps 8 and 11 refers to the sd-RPCA, and the steps 9 and 12 refers to the kd-RPCA. It is also important to observe that the anomaly detection from new observations does not require new robust subspace learning, it is only necessary to compute a moment-based Mahalanobis distance to classify the $c$ largest distances as anomalous.

@@

This section presents the performed experiments and the acquired results. First, in Section \ref{sec:AnalyzedScenario}, the experimental scenario adopted in the experiments is summarized. Then, Section \ref{sec:results} shows the results of the proposed approach and compares to selected anomaly detection algorithms.

@@

Anomaly detection algorithms usually rely on supervised or unsupervised methods, where the former requires labeled normal and anomalous data for training the model, while the latter does not require labeled data or training. However, the availability of labeled data is a challenging concern in real world problems of anomalous detection, where anomalies are rare or unkown events.

Semi-supervised algorithms are an alternative for the anomaly detection problem, considering that this method only relies on normal data for training and that non-malicious data can be obtained from historical information and from rule-based approaches. The proposed approach is semi-supervised and relies on normal data $\boldsymbol{X}^n$ and on contamination rate $c$ for training and anomalous detection, respectively.

Only for evaluation purpose, we use part of the contaminated data, denoted as $\textbf{X}^v$, for cross-validation, in order to estimate the contamination rate $c$ that best detect anomalies from $\textbf{X}^v$. However, we assume that the $c$ is well known or can be estimated for real world problems of anomaly detection.

Clustering methods for anomaly detection rely on three key assumptions: legitimate data instances often fall into a cluster whereas attacks do not, legitimate data instances are usually located near the closest cluster centroid while anomaly ones are often far away from it, legitimate data instances fall into vast and dense clusters and anomalies into small or spare ones \cite{ahmed2016survey, moustafa2019holistic}. Gaussian Mixture Model (GMM) is a statistical-based and parametric approach algorithm that estimates the distribution of the normal class from a training set and is typically based on a set of kernels \cite{moustafa2019holistic}.

This experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statistical approaches. As clustering-based method, we compare to K-means, which is the clustering algorithm most adopted for general purposes and for network anomaly detection \cite{gaddam2007kmeans,moustafa2019holistic}. It is observed that network traffic do not belong to a Gaussian distribution \cite{benson2010network,moustafa2019holistic}, and that algorithms based on non-Gaussian distributions, such as GMM, has been adopted for network anomaly detection \cite{moustafa2019holistic}. Therefore, we compare our proposal to GMM. We also evaluate the results of ROBPCA \cite{hubert2005robpca}, which is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.

In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of the CTU-13 data set individually, in a semi-supervised approach that does not rely on training data with labeled anomalies, in order to evaluate our proposed approach for all botnet malwares of the CTU-13. Therefore, the data set and scenarios for our evaluation can be denoted as $\boldsymbol{X} = \{\boldsymbol{X}_{10}, \boldsymbol{X}_{11}, ... , \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to Table \ref{tab:tab01}.

Given a contaminated data set $\boldsymbol{X}$, split it into $\boldsymbol{X}^n$ containing 60\% of the the normal data, and into cross validation $\boldsymbol{X}^v$, composed by 20\% of the normal and 50\% of the anomalous data, and into test $\boldsymbol{X}^t$ containing 20\% of the normal and 50\% of the anomalous data.

Evaluate if the code is correct and implement it as funcions in my code

try other distributions between train cv and test

Explain tha numbers of anomalies and normal for CV and test

@@

@@

@@

This paper evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection. Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.

@@

% Initially, the data $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ is split into normal data $\textbf{X}^r$ for robust subspace learning, and into $\textbf{X}^t$ for testing the anomaly detection on contaminated data, that can also be viewed as new observations of a production environment.
